---
title: Has the housing share really increased?
author: Osman Keshawarz
bibliography: prospectus.bib
output:
    tufte::tufte_handout: 
        number_sections: true
        fig_caption: true
        df_print: kable
---

```{r front_matter, echo=F}
library(csv)
library(tidyverse)
```


# Introduction

# Capital Share: Housing vs. Non-housing

@piketty2014's paper, focused mainly on wealth-income ratios, forms the basis
of Rognlie's decomposition of the capital share. This data is made available at
Zucman's website [@zucman]. Rognlie simply uses Zucman's decomposition of the
capital share, which I have reproduced in *Figure 1*. Though Rognlie's analysis
includes all G7 countries, the US trend closely follows the weighted trend from
the paper.

```{r fig1, fig.cap="Net Shares for Total Capital and Housing Capital Income of the US, 1950-2010", echo=F, fig.margin=T, fig.height=9.5, fig.width=5}
pzdata <- as.data.frame(read.csv("pzdata/pz_inc_50.csv"))
pzdata_obj <- ts(as.data.frame(read.csv("pzdata/pz_inc_50.csv")), start = 1950, end = 2010,
    freq = 1)
 
par(mfrow=c(2, 1))
plot.ts((pzdata_obj[, "capinc"]/pzdata_obj[, "natincome"]), col="blue", lwd=1.5, ylab="Net Capital Share", xlab="Year")
plot.ts(pzdata_obj[, "housing_capinc"]/pzdata_obj[, "natincome"], ylab="Net
Housing Capital Share", xlab="Year", col="blue", lwd=1.5)
```

```{r table1, fig.cap="Decadal Averages for Net Capital Share, 1950s-Present", echo=F}

pzdata <- pzdata %>%
            mutate(nonhousing_capinc = net_corp_capinc + self_capinc + net_for_capinc + net_govt_interest)

pz50 <- filter(pzdata, year <= 1959)
pz60 <- filter(pzdata, year %in% c(1960:1969))
pz70 <- filter(pzdata, year %in% c(1970:1969))
pz80 <- filter(pzdata, year %in% c(1980:1969))
pz90 <- filter(pzdata, year %in% c(1990:1969))
pz00 <- filter(pzdata, year %in% c(2000:2010))
row1 <- c(
            mean(pz50$housing_capinc/pz50$natincome),
            mean(pz60$housing_capinc/pz60$natincome),
            mean(pz70$housing_capinc/pz70$natincome),
            mean(pz80$housing_capinc/pz80$natincome),
            mean(pz90$housing_capinc/pz90$natincome),
            mean(pz00$housing_capinc/pz00$natincome))
            
row2 <- c(            
            mean((pz50$net_corp_capinc+pz50$self_capinc+pz50$net_for_capinc+pz50$net_govt_interest)/pz50$natincome),
            mean((pz60$net_corp_capinc+pz60$self_capinc+pz60$net_for_capinc+pz60$net_govt_interest)/pz60$natincome),
            mean((pz70$net_corp_capinc+pz70$self_capinc+pz70$net_for_capinc+pz70$net_govt_interest)/pz70$natincome),
            mean((pz80$net_corp_capinc+pz80$self_capinc+pz80$net_for_capinc+pz80$net_govt_interest)/pz80$natincome),
            mean((pz90$net_corp_capinc+pz90$self_capinc+pz90$net_for_capinc+pz90$net_govt_interest)/pz90$natincome),
            mean((pz00$net_corp_capinc+pz00$self_capinc+pz00$net_for_capinc+pz00$net_govt_interest)/pz00$natincome))

table1 <- rbind(row1, row2)
```

PZ's data on capital shares is extracted directly from the BEA's NIPA
tables^[specifically: NIPA 7.4.5 lines 13-15 & 16 and NIPA 1.3.5 line 11.].
Rognlie simply extracts the relevant data from PZ's data tables. In *Figures 1*
and *2*, I have replicated the result for the USA only. In *Figure 2*, the
rising trend in housing capital income vs. non-housing is small, but evident.
*Figure 3* displays the labor-capital shares directly, lifted again from PZ's
data set. It can be seen that factor income distributions remain relatively
constant, *a la* the Kaldor facts, up intil recently (the 1980s or so, which is
the same period in which the housing capital share begins to rise), when the
labor share begins its downward trend.

```{r fig4, fig.cap="Net Capital Shares of Corporate Value-Added, USA", echo=F, fig.height=4, fig.width=5}
par(mfrow=c(1,1))
plot((pzdata$housing_capinc/pzdata$natincome)~pzdata$year, type="l", col="blue", lwd=1.2,
ylim=c(0.02, 0.23), xlab="Year", ylab = "Net Capital Share")
lines((pzdata$nonhousing_capinc/pzdata$natincome)~pzdata$year, col="red", lwd=1.2)
legend(1950, 0.12, c("Housing", "Non-Housing"), col=c("blue", "red"), lwd=1)
```

```{r capshare, fig.cap="Factor Shares of National Income, USA 1950-2010", echo=F, fig.height=4, fig.width=5}
plot((pzdata$capinc/pzdata$natincome)~pzdata$year, type="l", col="blue", ylim=c(0.2, 0.8), xlab="Year", ylab="Share of National Income")
lines((pzdata$labinc/pzdata$natincome)~pzdata$year, col="red")
legend(1980, 0.6, c("Capital share", "Labor share"), col=c("blue", "red"), lwd=1)
``` 

\newpage

# Housing Share: Structures vs. Equipment vs. Land

Rognlie's other major finding (in the context of the rise of the housing
share) is displayed on Figure 6 [@rognlie2015; 22]. The figure decomposes
the housing share of national income into returns on structures, return on
equipment, return on land, and "pure profit." These figures are derived
from the US NIPA and Financial Accounts tables, which I will attempt to
replicate below.

```{r rog_fig_6, echo=FALSE, fig.cap="Net Capital Share"}

```

\newpage

# References




